

# AR, MA and ARMA Models

We now turn our attention to the auto-regressive and moving average model - the
second workhorse for time-series modelling.

## Building Blocks

These models are built from a number of elementary model types that we first
discuss before combining in different ways to model time series data seen in
the real world.


### White Noise Series

A time series $w_t$ is \emph{discrete white noise} if the $w_t$ are i.i.d with
a mean of zero. They all have the same variance $\sigma^2$ and
$\text{Cor}(w_i, w_j) = 0$ for $i \neq j$.

In addition, if $w_j \sim N(0, \sigma^2)$ then it is called
\emph{Gaussian white noise}.


A time series $x_t$ is a \emph{random walk} if

$$
x_t = x_{t-1} + w_t,
$$

where $w_t$ is a white-noise series.


```{r white_noise_series, echo=TRUE}
wn_innov <- rnorm(200, 0, 1)

sim_wn_ts <- arima.sim(list(),
                       n = length(wn_innov),
                       innov = wn_innov,
                       start.innov = rep(0, 100)
                      )

sim_wn_ts_tbl <- sim_wn_ts %>%
  tk_tbl(preserve_index = FALSE) %>%
  add_column(index = 1:length(sim_wn_ts), .before = 1)

ggplot(sim_wn_ts_tbl) +
  geom_line(aes(x = index, y = value)) +
  expand_limits(y = 0) +
  xlab("Index") +
  ylab("Value") +
  ggtitle("Plot of Generated Gaussian White Noise Series")
```

We also produce a standard and partial correlogram, but expect no significant
lags.

```{r white_noise_acf_pacf, echo=TRUE}
sim_wn_ts %>% Acf()  %>% autoplot() + ggtitle("ACF for White Noise Series")
sim_wn_ts %>% Pacf() %>% autoplot() + ggtitle("PACF for White Noise Series")
```


### Autoregressive Process - AR(p)

The time series $x_t$ is an \emph{auto-regressive process of order $p$},
$\text{AR}(p)$, if,

\begin{eqnarray*}
x_t &=& \delta + \alpha_1 x_1 + ... + \alpha_p x_p + w_t \\
    &=& \alpha_0 + \sum^p_{i=1} \alpha_i x_{t-i} + w_t,
\end{eqnarray*}

where

\begin{eqnarray*}
w_t      &=&    \text{is a white-noise process} \\
\delta   &=&    \text{'intercept' value}     \\
\alpha_p &\neq& 0 \text{ for order } p
\end{eqnarray*}


With this definition, we can use this as a *data-generating process* to
construct an example AR(1) model, and visual inspect the results.

```{r generate_ar1_model, echo=TRUE}
sim_ar1_ts <- arima.sim(list(ar = 0.5),
                        n     = length(wn_innov),
                        innov = wn_innov,
                        start.innov = rep(0, 100)
                        )

sim_ar1_ts_tbl <- sim_ar1_ts %>%
  tk_tbl(preserve_index = FALSE) %>%
  add_column(index = 1:length(sim_ar1_ts), .before = 1)

ggplot(sim_ar1_ts_tbl) +
  geom_line(aes(x = index, y = value)) +
  expand_limits(y = 0) +
  xlab("Index") +
  ylab("Value") +
  ggtitle("Plot of Generated AR(1) Series")
```

Having generated the AR(1) process, we now look at the correlogram to see what
it produces.

```{r ar1_acf_pacf, echo=TRUE}
sim_ar1_ts %>% Acf()  %>% autoplot() + ggtitle("ACF for AR(1) Series")
sim_ar1_ts %>% Pacf() %>% autoplot() + ggtitle("PACF for AR(1) Series")
```


Moving on, in a similar fashion we can construct an AR(2) model with the same
innovations, but with $\alpha_1 = 1$, $\alpha_2 = -0.25$:


```{r generate_ar2_model, echo=TRUE}
sim_ar2_ts <- arima.sim(list(ar = c(1, -0.25)),
                        n     = length(wn_innov),
                        innov = wn_innov,
                        start.innov = rep(0, 100)
                         )

sim_ar2_ts_tbl <- sim_ar2_ts %>%
  tk_tbl(preserve_index = FALSE) %>%
  add_column(index = 1:length(sim_ar2_ts), .before = 1)


ggplot(sim_ar2_ts_tbl) +
  geom_line(aes(x = index, y = value)) +
  expand_limits(y = 0) +
  xlab("Index") +
  ylab("Value") +
  ggtitle("Plot of Generated AR(2) Series")
```

We also look at the correlogram for this series.

```{r ar2_acf_pacf, echo=TRUE}
sim_ar2_ts %>% Acf()  %>% autoplot() + ggtitle("ACF for AR(2) Series")
sim_ar2_ts %>% Pacf() %>% autoplot() + ggtitle("PACF for AR(2) Series")
```



We can try a number of different parameters here to see the effect - we try
another AR(2) but this time with both parameters positive:


```{r generate_ar2alt_model, echo=TRUE}
sim_ar2alt_ts <- arima.sim(list(ar = c(0.5, 0.4)),
                           n     = length(wn_innov),
                           innov = wn_innov,
                           start.innov = rep(0, 100)
                           )

sim_ar2alt_ts_tbl <- sim_ar2alt_ts %>%
  tk_tbl(preserve_index = FALSE) %>%
  add_column(index = 1:length(sim_ar2alt_ts), .before = 1)


ggplot(sim_ar2alt_ts_tbl) +
  geom_line(aes(x = index, y = value)) +
  expand_limits(y = 0) +
  xlab("Index") +
  ylab("Value") +
  ggtitle("Plot of Generated Alternate AR(2) Series")
```

And the correlogram for this alternate AR(2) series.

```{r ar2alt_acf_pacf, echo=TRUE}
sim_ar2alt_ts %>% Acf()  %>% autoplot() + ggtitle("ACF for Alternate AR(2) Series")
sim_ar2alt_ts %>% Pacf() %>% autoplot() + ggtitle("ACF for Alternate AR(2) Series")
```

To help with intuition we plot these generated data together.

```{r plot_generated_ar_series, echo=TRUE}
sim_compare_tbl <- list(
  wn     = sim_wn_ts_tbl,
  ar1    = sim_ar1_ts_tbl,
  ar2    = sim_ar2_ts_tbl,
  ar2alt = sim_ar2alt_ts_tbl
  ) %>%
  bind_rows(.id = 'model')


ggplot(sim_compare_tbl) +
  geom_line(aes(x = index, y = value, colour = model)) +
  xlab("Index") +
  ylab("Value") +
  ggtitle("Generated AR Models Comparison Plot")
```


### Moving Average Process - MA(q)

A moving average (MA) process of order $q$ is a linear combination of the
current white noise term and the $q$ most recent past white noise terms,

\begin{eqnarray*}
x_t &=& w_t + \gamma + \beta_1 w_1 + ... + \beta_q w_q \\
    &=& \gamma + w_t + \sum^q_{i=1} \beta_i w_{t-i}
\end{eqnarray*}

where $w_t$ is a white-noise process with mean 0 and variance $\sigma^2$.

\begin{eqnarray*}
w_t     &=&    \text{is a white-noise process with mean 0 and variance } \sigma^2 \\
\gamma  &=&    \text{'intercept' value}     \\
\beta_q &\neq& 0 \text{ for order } q
\end{eqnarray*}


```{r generate_ma1_model, echo=TRUE}
sim_ma1_ts <- arima.sim(list(ma = 0.5),
                        n     = length(wn_innov),
                        innov = wn_innov,
                        start.innov = rep(0, 100)
                        )

sim_ma1_ts_tbl <- sim_ma1_ts %>%
  tk_tbl(preserve_index = FALSE) %>%
  add_column(index = 1:length(sim_ma1_ts), .before = 1)

ggplot(sim_ma1_ts_tbl) +
  geom_line(aes(x = index, y = value)) +
  expand_limits(y = 0) +
  xlab("Index") +
  ylab("Value") +
  ggtitle("Plot of Generated MA(1) Series")
```

We check the MA(1) correlogram:

```{r ma1_acf_pacf, echo=TRUE}
sim_ma1_ts %>% Acf()  %>% autoplot() + ggtitle("ACF for MA(1) Series")
sim_ma1_ts %>% Pacf() %>% autoplot() + ggtitle("ACF for MA(1) Series")
```


Like we did before, we now generate an MA(2) model.

```{r generate_ma2_model, echo=TRUE}
sim_ma2_ts <- arima.sim(list(ma = c(1, -0.25)),
                        n     = length(wn_innov),
                        innov = wn_innov,
                        start.innov = rep(0, 100)
                         )

sim_ma2_ts_tbl <- sim_ma2_ts %>%
  tk_tbl(preserve_index = FALSE) %>%
  add_column(index = 1:length(sim_ma2_ts), .before = 1)


ggplot(sim_ma2_ts_tbl) +
  geom_line(aes(x = index, y = value)) +
  expand_limits(y = 0) +
  xlab("Index") +
  ylab("Value") +
  ggtitle("Plot of Generated MA(2) Series")
```

Once again we look at the MA(2) series.

```{r ma2_acf_pacf, echo=TRUE}
sim_ma2_ts %>% Acf()  %>% autoplot() + ggtitle("ACF for MA(2) Series")
sim_ma2_ts %>% Pacf() %>% autoplot() + ggtitle("ACF for MA(2) Series")
```


```{r plot_generated_ma_series, echo=TRUE}
sim_compare_tbl <- list(
  wn  = sim_wn_ts_tbl,
  ma1 = sim_ma1_ts_tbl,
  ma2 = sim_ma2_ts_tbl
  ) %>%
  bind_rows(.id = 'model')


ggplot(sim_compare_tbl) +
  geom_line(aes(x = index, y = value, colour = model)) +
  xlab("Index") +
  ylab("Value") +
  ggtitle("Generated MA Models Comparison Plot")
```


### Comparing AR and MA Models

Before moving on to ARMA and ARIMA models we combine all our generated series
and plot them all against each other.

```{r plot_generated_ar_ma_series, echo=TRUE}
sim_compare_tbl <- list(
  wn     = sim_wn_ts_tbl,
  ar1    = sim_ar1_ts_tbl,
  ar2    = sim_ar2_ts_tbl,
  ar2alt = sim_ar2alt_ts_tbl,
  ma1    = sim_ma1_ts_tbl,
  ma2    = sim_ma2_ts_tbl
  ) %>%
  bind_rows(.id = 'model')


ggplot(sim_compare_tbl) +
  geom_line(aes(x = index, y = value, colour = model)) +
  xlab("Index") +
  ylab("Value") +
  ggtitle("Generated AR and MA Models Comparison Plot")
```
