
## Automated Model Selection

The `ets()` function allows us to automatically select a model given the time
series data.

The exponentially-smoothed models fitted allows us to use AIC and BIC as a
decision metric.

The Aikake Information Criterion (AIC) is defined to be:

$$
\text{AIC} = -2 \log(L) + 2k,
$$

where $L$ is the likelihood of the fitted model and $k$ is the number of
degrees of freedom in the model (free parameters, initial states and residual
variance).

For small sample sizes this approach may still be prone to overfitting, so we
may use a corrected version for small-sample bias, ($\text{AIC}_c$)

$$
\text{AIC}_c = \text{AIC} + \frac{k(k+1)}{T - k -1},
$$

The Bayes Information Criterion (BIC) is defined to be:

$$
\text{BIC} = \text{AIC} + k (\log(T) - 2)
$$

The `ets()` model allows us to select the type of error term, trend type and
seasonality, letting us choose additive (A), multiplicative (M), none (N) or
automatically selected (Z) for each of the three options.

The first values, the error term, cannot be 'N' and a number of combinations
are numerically unstable and suppressed.

We now fit both the Air Passenger and Maine Unemployment data using the
automatic model selection.

### Air Passenger Models

We now fit the Air Passenger data first.

```{r fit_ap_auto_ets, echo=TRUE}
ap_autoets_fit <- ap_train_tbl %>%
  tk_ts(select = value, start = 1949, frequency = 12) %>%
  ets(model = 'ZZZ')

ap_autoets_forecast <- ap_autoets_fit %>%
  forecast(h = 12, level = c(50, 80))

ap_autoets_forecast_tbl <- ap_autoets_forecast %>%
  sw_sweep(rename_index = 'month')

ap_autoets_fit %>% sw_tidy()
ap_autoets_fit %>% sw_glance()

ggplot(ap_autoets_forecast_tbl) +
  geom_line(aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = ap_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Air Passengers Forecast for Automatic ETS Model')
```

Having fit the data and selected the model, we now can compare our different
models and how the forecasts compare to each other.

```{r compare_ap_forecasts, echo=TRUE}
ap_forecast_compare_tbl <- list(
  ewma   = ap_ewma_forecast_tbl    %>% filter(key == 'forecast'),
  hw     = ap_hw_forecast_tbl      %>% filter(key == 'forecast'),
  hwmult = ap_hwmult_forecast_tbl  %>% filter(key == 'forecast'),
  auto   = ap_autoets_forecast_tbl %>% filter(key == 'forecast')
  ) %>%
  bind_rows(.id = 'model') %>%
  mutate(month = as.yearmon(month))


ggplot(ap_forecast_compare_tbl) +
  geom_line(aes(x = month, y = value, colour = model)) +
  geom_line(aes(x = month, y = value), colour = 'black', data = ap_test_tbl) +
  expand_limits(y = 0) +
  scale_x_yearmon() +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Comparison of Air Passenger Forecast Models')
```


### Maine Unemployment Models

We now move on to the Maine Unemployment data, and repeat the procedure.


```{r fit_maine_auto_ets, echo=TRUE}
maine_autoets_fit <- maine_train_tbl %>%
  tk_ts(select = value, start = 1996, frequency = 12) %>%
  ets(model = 'ZZZ')

maine_autoets_forecast <- maine_autoets_fit %>%
  forecast(h = 20, level = c(50, 80))

maine_autoets_forecast_tbl <- maine_autoets_forecast %>%
  sw_sweep(rename_index = 'month')


maine_autoets_fit %>% sw_tidy()
maine_autoets_fit %>% sw_glance()


ggplot(maine_autoets_forecast_tbl) +
  geom_line(aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = maine_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Maine Unemployment Forecast for Automatic ETS Model')
```

As before, we compare the forecasts against one another.

```{r compare_maine_forecasts, echo=TRUE}
maine_forecast_compare_tbl <- list(
  ewma      = maine_ewma_forecast_tbl          %>% filter(key == 'forecast'),
  ewmaparam = maine_ewma_paramfit_forecast_tbl %>% filter(key == 'forecast'),
  hw        = maine_hw_forecast_tbl            %>% filter(key == 'forecast'),
  hwmult    = maine_hwmult_forecast_tbl        %>% filter(key == 'forecast'),
  auto      = maine_autoets_forecast_tbl       %>% filter(key == 'forecast')
  ) %>%
  bind_rows(.id = 'model') %>%
  mutate(month = as.yearmon(month))


ggplot(maine_forecast_compare_tbl) +
  geom_line(aes(x = month, y = value, colour = model)) +
  geom_line(aes(x = month, y = value), colour = 'black', data = maine_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Comparison of Maine Unemployment Forecast Models')
```


# Forecast Evaluation

We finally come to the most important part of forecasting - how we decide how
good it is.


## Evaluation Metrics

We start by defining the forecast *error* or forecast *residual*, $e_{t+h}$ as:

$$
e_{t+h} = y_{t+h} - \hat{y}_{t+h \, | \,t}
$$

This is scale-dependent, so we also define the *forecast percentage error*,
$p_t$

$$
p_t = \frac{e_t}{y_t}
$$


where:

\begin{eqnarray*}
y_{t+h} &=& \text{realized value at time } t+h   \\
\hat{y}_{t+h} &=& \text{forecast value at time } t+h
\end{eqnarray*}

From this we build a number of different metrics:

\begin{eqnarray*}
\text{Mean absolute error, MAE}              &=&
    \text{mean}(|e_t|)                                     \\
\text{Root mean square error, RMSE}          &=&
    \sqrt{\text{mean}(e_t^2)}                              \\
\text{Mean absolute percentage error, MAPE}  &=& 
    \text{mean}(|p_t|)                                     \\
\text{Symmetric MAPE}                        &=&
    \frac{200 * |y_t - \hat{y}_t|}{y_t + \hat{y}_t}        \\
\text{Scaled error}, q_j                     &=&
    \frac{e_j}{\frac{1}{T-1} \sum_{t=2}^T |y_t - y_{t-1}|} \\
\text{Mean absolute scaled error, MASE}      &=&
    \text{mean}(|q_j|)
\end{eqnarray*}


## Evaluating Fits

We run the accuracy metrics for both the Air Passenger and the Maine
Unemployment data for the fits we ran.


### Air Passenger Metrics

We now calculate the evaluation metrics for the Air Passenger models.

```{r calculate_ap_metrics, echo=TRUE}
ap_ets_assess_tbl <- list(
  ewma    = ap_ewma_forecast,
  hw      = ap_hw_forecast,
  hwmult  = ap_hwmult_forecast,
  autoets = ap_autoets_forecast
  ) %>%
  map(accuracy, x = ap_test_tbl$value) %>%
  map(as_tibble, rownames = 'type') %>%
  bind_rows(.id = 'model') %>%
  arrange(type, model)

ap_ets_assess_tbl %>% summary()
```

We plot the values.

```{r plot_ap_metrics, echo=TRUE}
plot_tbl <- ap_ets_assess_tbl %>%
  gather('metric', 'value', -model, -type)
  
ggplot(plot_tbl) +
  geom_col(aes(x = model, y = value, fill = type), position = 'dodge') +
  facet_wrap(~ metric, scales = 'free_y') +
  xlab('Model') +
  ylab('Metric') +
  ggtitle('Facet Plot for Air Passenger Forecast Metrics')
```



### Maine Unemployment Metrics

We repeat this process for the Maine Unemployment series.

```{r calculate_maine_metrics, echo=TRUE}
maine_ets_assess_tbl <- list(
  ewma    = maine_ewma_forecast,
  ewma    = maine_ewma_paramfit_forecast,
  hw      = maine_hw_forecast,
  hwmult  = maine_hwmult_forecast,
  autoets = maine_autoets_forecast
  ) %>%
  map(accuracy, x = maine_test_tbl$value) %>%
  map(as_tibble, rownames = 'type') %>%
  bind_rows(.id = 'model') %>%
  arrange(type, model)

maine_ets_assess_tbl %>% summary()
```

We plot the values.

```{r plot_maine_metrics, echo=TRUE}
plot_tbl <- maine_ets_assess_tbl %>%
  gather('metric', 'value', -model, -type)
  
ggplot(plot_tbl) +
  geom_col(aes(x = model, y = value, fill = type), position = 'dodge') +
  facet_wrap(~ metric, scales = 'free_y') +
  xlab('Model') +
  ylab('Metric') +
  ggtitle('Facet Plot for Maine Unemployment Forecast Metrics')
```
