
## Forecasting with Linear Models
============================================
column {.tabset}
--------------------
### Forecasting
<h4> Forecating</h4>
One possible starting point for forecasting problems is to use a linear model,
much like any other predictive modelling with time as an input to the
regression. We account for seasonality by using the various calendar labels as
additional predictors.

Rather than fit the model on the whole dataset, we use the final year of data
as our test dataset.

```{r ap_forecast_regression_data, echo=TRUE}
ap_lm_data_tbl <- airpassengers_tbl %>%
  mutate(month_label = format(month, '%m'),
         year_num    = year(month)
         )

ap_train_tbl <- ap_lm_data_tbl %>% filter(year_num != 1960)
ap_test_tbl  <- ap_lm_data_tbl %>% filter(year_num == 1960)
```

```{r maine_forecast_regression_data, echo=TRUE}
maine_train_tbl <- maine_tbl %>% filter(month <  as.yearmon('2005-01'))
maine_test_tbl  <- maine_tbl %>% filter(month >= as.yearmon('2005-01'))
```



### Simple Linear Model

We now fit a model.

```{r ap_forecast_regression_lm, echo=TRUE}
ap_base_lm <- lm(value ~ 0 + month_label + year_num, data = ap_train_tbl)
```
```{r}
ap_base_lm %>% glance()
ap_base_lm %>% tidy()
```

We now check the fitted values against the observed.

```{r ap_forecast_regression_train, echo=TRUE}
plot_tbl <- ap_base_lm %>%
  augment() %>%
  mutate(month = paste(year_num, month_label, sep = '-') %>% as.yearmon()) %>%
  select(month, observed = value, fitted = .fitted) %>%
  gather('series', 'value', -month)

ggplot(plot_tbl) +
  geom_line(aes(x = month, y = value, colour = series)) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle("Comparison Lineplot of Air Passengers vs Linear Model")
```

We see that the additive model here may be less appropriate as a multiplicative
model as our linear model over-estimates early values and then under-estimates
towards the end of the series.

We expect this will continue when we try the model with the test data.

```{r ap_forecast_regression_test, echo=TRUE}
plot_tbl <- ap_base_lm %>%
  augment(newdata = ap_test_tbl) %>%
  mutate(month = paste(year_num, month_label, sep = '-') %>% as.yearmon()) %>%
  select(month, observed = value, fitted = .fitted) %>%
  gather('series', 'value', -month)

ggplot(plot_tbl) +
  geom_line(aes(x = month, y = value, colour = series)) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle("Comparison Lineplot of Air Passengers vs Linear Model (1960 points)")
```


### Poisson Model

A more appropriate model may be to fit the Poisson model.

```{r ap_forecast_regression_glm, echo=TRUE}
ap_base_glm <- glm(value ~ 0 + month_label + year_num,
                   data   = ap_train_tbl,
                   family = poisson)

ap_base_glm %>% glance()
ap_base_glm %>% tidy()
```

We now check the fitted values against the observed.

```{r ap_forecast_regression_glm_train, echo=TRUE}
plot_tbl <- ap_base_glm %>%
  augment(type.predict = 'response') %>%
  mutate(month = paste(year_num, month_label, sep = '-') %>% as.yearmon()) %>%
  select(month, observed = value, fitted = .fitted) %>%
  gather('series', 'value', -month)

ggplot(plot_tbl) +
  geom_line(aes(x = month, y = value, colour = series)) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle("Comparison Lineplot of Air Passengers vs Poisson Model")
```

We see that a Poisson model does a better job of capturing the magnitude of
the values throughout the series. We hope this means the 1960 data points are
better matched also.

```{r ap_forecast_regression_glm_test, echo=TRUE}
plot_tbl <- ap_base_glm %>%
  augment(newdata = ap_test_tbl, type.predict = 'response') %>%
  mutate(month = paste(year_num, month_label, sep = '-') %>% as.yearmon()) %>%
  select(month, observed = value, fitted = .fitted) %>%
  gather('series', 'value', -month)

ggplot(plot_tbl) +
  geom_line(aes(x = month, y = value, colour = series)) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle("Comparison Lineplot of Air Passengers vs Poisson Model (1960 points)")
```
