


## ARMA and ARIMA Models

Autoregressive moving-average models combine both the AR and MA models
together. The benefit of combining both models is due to the
*principle of parsimony* - given two equal quality models, we prefer the one
with the lower number of parameters.

In general, ARMA models fit time series with a lower number of parameters than
either AR or MA models alone, so they are preferred.

More formally, we define an $\text{ARMA}(p, q)$ model to be

$$
x_t = \sum_{i=1}^p \alpha_i x_{t-i} + w_t + \sum_{j=1}^q \beta_j w_{t-j}
$$

where $w_t$ is white noise.

Both $\text{AR}(p)$ and $\text{MA}(q)$ models are special cases of
$\text{ARMA}(p, q)$ (with $q = 0$ and $p = 0$ respectively).

We now fit an $\text{ARMA}(1, 1)$ model using the same white-noise innovations
as before.


```{r generate_arma_1_1_model, echo=TRUE}
sim_arma_1_1_ts <- arima.sim(list(ar = 0.5, ma = 0.5),
                             n     = length(wn_innov),
                             innov = wn_innov,
                             start.innov = rep(0, 100)
                             )

sim_arma_1_1_ts_tbl <- sim_arma_1_1_ts %>%
  tk_tbl(preserve_index = FALSE) %>%
  add_column(index = 1:length(sim_arma_1_1_ts), .before = 1)

ggplot(sim_arma_1_1_ts_tbl) +
  geom_line(aes(x = index, y = value)) +
  expand_limits(y = 0) +
  xlab("Index") +
  ylab("Value") +
  ggtitle("Plot of Generated ARMA(1,1) Series")
```

We use `auto.arima()` to fit parameters for this data to see if we recover the
given parameters.

```{r fit_sim_arma_1_1, echo=TRUE}
sim_arma_1_1_ts %>%
  auto.arima() %>%
  sw_tidy()
```


### ARIMA Models

As discussed earlier, it is often more effective to model the differences in
a time series rather than the raw values.

A variant of the ARMA is the *Auto-Regressive Integrated Moving Average* model
which applies the ARMA model to the differences. ARIMA model are specified With
a third order value, $d$ - the order of the differences, and so are described
as $\text{ARIMA}(p, d, q)$ models.

While there is no theoretical limit to $d$, in practice it is uncommon to see
$d > 1$.


## Fitting ARMA and ARIMA Models

We now turn our attention to fitting ARMA and ARIMA models.

In some cases, we use a combination of domain knowledge and autocorrelation
analysis to determine the order of the AR(I)MA model we fit, but generally we
go with the best fit as determined by AIC or BIC (and ensure this 'best' fit
is still useful and makes sense).


### Fitting the Air Passenger Data

As always, we start with the Air Passenger data.

Purely for illustration, we start with the ARIMA(1, 0, 1) model - which is
equivalent to the ARMA(1, 1) - and then try the ARIMA(1, 1, 1).

```{r fit_ap_arma_1_1_arima, echo=TRUE}
ap_arma_1_1_fit <- ap_train_tbl %>%
  tk_ts(select = value, start = 1949, frequency = 12) %>%
  Arima(order = c(1, 0, 1))

ap_arma_1_1_forecast <- ap_arma_1_1_fit %>%
  forecast(h = 12, level = c(50, 80))

ap_arma_1_1_forecast_tbl <- ap_arma_1_1_forecast %>%
  sw_sweep(rename_index = 'month')


ap_arma_1_1_fit %>% sw_tidy()
ap_arma_1_1_fit %>% sw_glance()


ggplot(ap_arma_1_1_forecast_tbl) +
  geom_line(aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = ap_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Air Passenger Forecast for ARMA(1, 1) Model')
```

This is not a good model.

```{r fit_ap_arma_1_1_1_arima, echo=TRUE}
ap_arma_1_1_1_fit <- ap_train_tbl %>%
  tk_ts(select = value, start = 1949, frequency = 12) %>%
  Arima(order = c(1, 1, 1))

ap_arma_1_1_1_forecast <- ap_arma_1_1_1_fit %>%
  forecast(h = 12, level = c(50, 80))

ap_arma_1_1_1_forecast_tbl <- ap_arma_1_1_1_forecast %>%
  sw_sweep(rename_index = 'month')


ap_arma_1_1_1_fit %>% sw_tidy()
ap_arma_1_1_1_fit %>% sw_glance()


ggplot(ap_arma_1_1_1_forecast_tbl) +
  geom_line(aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = ap_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Air Passenger Forecast for ARIMA(1, 1, 1) Model')
```

This is not a good model either. We need to include a seasonal component to
this process to use it properly.

Before we move on, we try using the `auto.arima()` function to see what the
best model looks like.

```{r ap_autoarima_fit, echo=TRUE}
ap_autoarima_fit <- ap_train_tbl %>%
  tk_ts(select = value, start = 1949, frequency = 12) %>%
  auto.arima(seasonal = FALSE)

ap_autoarima_forecast <- ap_autoarima_fit %>%
  forecast(h = 12, level = c(50, 80))

ap_autoarima_forecast_tbl <- ap_autoarima_forecast %>%
  sw_sweep(rename_index = 'month')


ap_autoarima_fit %>% sw_tidy()
ap_autoarima_fit %>% sw_glance()


ggplot(ap_autoarima_forecast_tbl) +
  geom_line(aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = ap_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Air Passenger Forecast for Auto-ARIMA Model')
```

We will look at including seasonal elements to the ARIMA model shortly.


### Fitting the Maine Unemployment Data

We will not try to fit any particular order of ARIMA model for the Maine
Unemployment data and move straight to using `auto.arima()` to fit the models.

```{r maine_autoarima_fit, echo=TRUE}
maine_autoarima_fit <- maine_train_tbl %>%
  tk_ts(select = value, start = 1996, frequency = 12) %>%
  auto.arima(seasonal = FALSE)

maine_autoarima_forecast <- maine_autoarima_fit %>%
  forecast(h = 20, level = c(50, 80))

maine_autoarima_forecast_tbl <- maine_autoarima_forecast %>%
  sw_sweep(rename_index = 'month')

maine_autoarima_fit %>% sw_tidy()
maine_autoarima_fit %>% sw_glance()


ggplot(maine_autoarima_forecast_tbl) +
  geom_line(aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = maine_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Maine Unemployment Forecast for Auto-ARIMA Model')
```


## Seasonal ARIMA Models

We add seasonality to an ARIMA model by adding separate analysis for the
seasonal component of the time series.

Full detail of this approach is beyond the scope of this workshop, but we
specify a SARIMA model as $\text{SARIMA}(p, d, q)(P, D, Q)_m$, where the capitals
$(P, D, Q)$ are for the seasonal part of the model, with $m$ denoting the
number of data points per seasonal cycle - so we have $m = 12$ for monthly
data with annual seasonality.

### Fitting SARIMA Models

Fitting a Seasonal-ARIMA is a matter of adding the `seasonal = TRUE` parameter
to `auto.arima()`.

We now fit SARIMA models for both the Air Passenger and Maine Unemployment
data.

```{r ap_sarima_fit, echo=TRUE}
ap_sarima_fit <- ap_train_tbl %>%
  tk_ts(select = value, start = 1949, frequency = 12) %>%
  auto.arima(seasonal = TRUE)

ap_sarima_forecast <- ap_sarima_fit %>%
  forecast(h = 12, level = c(50, 80))

ap_sarima_forecast_tbl <- ap_sarima_forecast %>%
  sw_sweep(rename_index = 'month')


ap_sarima_fit %>% sw_tidy()
ap_sarima_fit %>% sw_glance()


ggplot(ap_sarima_forecast_tbl) +
  geom_line(aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = ap_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Air Passenger Forecast for Seasonal ARIMA Model')
```


Now we fit the Maine Unemployment data.


```{r maine_sarima_fit, echo=TRUE}
maine_sarima_fit <- maine_train_tbl %>%
  tk_ts(select = value, start = 1996, frequency = 12) %>%
  auto.arima(seasonal = TRUE)

maine_sarima_forecast <- maine_sarima_fit %>%
  forecast(h = 20, level = c(50, 80))

maine_sarima_forecast_tbl <- maine_sarima_forecast %>%
  sw_sweep(rename_index = 'month')


maine_sarima_fit %>% sw_tidy()
maine_sarima_fit %>% sw_glance()


ggplot(maine_sarima_forecast_tbl) +
  geom_line(aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = maine_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Maine Unemployment Forecast for Seasonal ARIMA Model')
```



## Forecast Evaluation

We now run forecast evaluations metrics for all our forecasts.


### Air Passenger Forecasts

We run the data for the Air Passenger data.

```{r calculate_ap_evaluations, echo=TRUE}
ap_metrics_tbl <- ls(pattern = '^ap_.*_forecast$') %>%
  enframe(value = 'variable') %>%
  mutate(model    = str_replace(variable, 'ap_(.*)_forecast', '\\1'),
         data     = map(variable, get),
         accuracy = map(data, accuracy, x = ap_test_tbl$value),
         acc_tbl  = map(accuracy, as_tibble, rownames = 'type')) %>%
  select(model, acc_tbl) %>%
  unnest() %>%
  arrange(type, model)

ap_metrics_tbl %>% summary()
```


```{r plot_full_ap_metrics, echo=TRUE}
plot_tbl <- ap_metrics_tbl %>%
  gather('metric', 'value', -model, -type)
  
ggplot(plot_tbl) +
  geom_col(aes(x = model, y = value, fill = type), position = 'dodge') +
  facet_wrap(~ metric, scales = 'free_y') +
  xlab('Model') +
  ylab('Metric') +
  ggtitle('Facet Plot for Air Passenger Forecast Metrics') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```


### Maine Unemployment Forecasts

And we run the data for the Maine Unemployment time series.

```{r calculate_maine_evaluations, echo=TRUE}
maine_metrics_tbl <- ls(pattern = '^maine_.*_forecast$') %>%
  enframe(value = 'variable') %>%
  mutate(model    = str_replace(variable, 'maine_(.*)_forecast', '\\1'),
         data     = map(variable, get),
         accuracy = map(data, accuracy, x = maine_test_tbl$value),
         acc_tbl  = map(accuracy, as_tibble, rownames = 'type')) %>%
  select(model, acc_tbl) %>%
  unnest() %>%
  arrange(type, model)

maine_metrics_tbl %>% summary()
```

```{r plot_full_maine_metrics, echo=TRUE}
plot_tbl <- maine_metrics_tbl %>%
  gather('metric', 'value', -model, -type)
  
ggplot(plot_tbl) +
  geom_col(aes(x = model, y = value, fill = type), position = 'dodge') +
  facet_wrap(~ metric, scales = 'free_y') +
  xlab('Model') +
  ylab('Metric') +
  ggtitle('Facet Plot for Maine Unemployment Forecast Metrics') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

Addtional Information
====================================

### R Environment

```{r show_session_info, echo=TRUE, message=TRUE}
devtools::session_info()
```
