

## Time Series Forecasting

Our main objective in forecasting is to estimate the value of a future
quantity, $x_{n+k}$, given past values ${x_1, x_2, ..., x_n}$. We assume no
seasonal or trend effects, or that such effects are removed from the data.

We assume that the underlying mean of the data is $\mu_t$, and that the value
$x_t$ changes from timestep to timestep, but this change is random.

Our model can be expressed as

$$
x_t = \mu_t + w_t,
$$

where

\begin{eqnarray*}
\mu_t &=&     \text{non-stationary mean of the process at time } t \\
w_t   &\sim&  \mathcal{N}(0, \sigma) \text{ (i.i.d)}
\end{eqnarray*}


### EWMA Models

We let $a_t$ be our estimate of $\mu_t$, and can define the
*exponentially-weighted moving average (EWMA)*, $a_t$ to be

$$
a_t = \alpha x_t + (1 - \alpha) \, a_{t-1}, \;\;\; 0 \leq \alpha \leq 1.
$$

The value of $\alpha$ controls the amount of smoothing, as is referred
to as the *smoothing parameter*. An alternative name for this time-series model
is *simple exponential smoothing (SES)*.

The main package we use for forecasting in R is the `forecast` package,
created by Rob Hyndman. `forecast` contains a convenient framework for
performing most of the task required for time series forecast. We then use
the `sweep` package to bring the outputs into a tidy format where appropriate.

#### Air Passenger Data

We start using the EWMA model on the Air Passenger data.

As a simple approach for testing our models, we remove the final year from the
air passenger data and produce forecasts for those, comparing them to the 
actual values.

```{r ap_ewma_first, echo=TRUE}
ap_ewma_fit <- ap_train_tbl %>%
  tk_ts(select = value, start = 1949, frequency = 12) %>%
  ets(model = 'ANN')

ap_ewma_fit %>% sw_tidy()
ap_ewma_fit %>% sw_glance()
```

Now that we have fit our model we can use the `forecast()` function to produce
forecasts for future times, as well as plotting them.

```{r ap_produce_forecasts, echo=TRUE}
ap_ewma_forecast <- ap_ewma_fit %>%
  forecast(h = 12, level = c(50, 80))

ap_ewma_forecast_tbl <- ap_ewma_forecast %>%
  sw_sweep(rename_index = 'month')


ggplot(ap_ewma_forecast_tbl) +
  geom_line  (aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = ap_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Forecast for Air Passengers with EWMA Model')
```

We see that the EWMA model is not effective in terms prediction.

We can compare the fits against the data directly with `sw_augment()`:

```{r ap_ewma_sweep_augment, echo=TRUE}
ap_ewma_augment_tbl <- ap_ewma_fit %>%
  sw_augment(rename_index = 'month')


ggplot(ap_ewma_augment_tbl) +
  geom_line(aes(x = month, y = .actual)) +
  geom_line(aes(x = month, y = .fitted), colour = 'red') +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Forecast for Air Passengers with EWMA Model')
```




#### Maine Unemployment

We now try the EWMA model on the Maine Unemployment data to see if this model
works any better.

```{r maine_ewma_model, echo=TRUE}
maine_ewma_fit <- maine_train_tbl %>%
  tk_ts(select = value, start = 1996, frequency = 12) %>%
  ets(model = 'ANN')

maine_ewma_fit %>% sw_tidy()
maine_ewma_fit %>% sw_glance()
```

As before, having fit the EWMA model we now produce forecasts.

```{r maine_produce_forecasts, echo=TRUE}
maine_ewma_forecast <- maine_ewma_fit %>%
  forecast(h = 20, level = c(50, 80))

maine_ewma_forecast_tbl <- maine_ewma_forecast %>%
  sw_sweep(rename_index = 'month')


ggplot(maine_ewma_forecast_tbl) +
  geom_line  (aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = maine_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Forecast for Maine Unemployment Data with EWMA Model')
```

In some cases, with such a simple model we may prefer to set values for
$\alpha$ directly rather than fitting it to the data. We do this by passing
the chosen parameters directly into the fitting routine:

```{r maine_ewma_model_set_params, echo=TRUE}
maine_ewma_paramfit <- maine_train_tbl %>%
  tk_ts(select = value, start = 1996, frequency = 12) %>%
  ets(model = 'ANN', alpha = 0.8)

maine_ewma_paramfit %>% sw_tidy()
maine_ewma_paramfit %>% sw_glance()
```

We now produce forecasts using the set EWMA parameter to see how they compare.

```{r maine_produce_paramfit_forecasts, echo=TRUE}
maine_ewma_paramfit_forecast <- maine_ewma_paramfit %>%
  forecast(h = 20, level = c(50, 80))

maine_ewma_paramfit_forecast_tbl <- maine_ewma_paramfit_forecast %>%
  sw_sweep(rename_index = 'month')


ggplot(maine_ewma_paramfit_forecast_tbl) +
  geom_line  (aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = maine_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Forecast for Maine Unemployment Data with Non-Fitted EWMA Model')
```


### Holt-Winters Model

The Holt-Winters method generalises this concept, allowing for trends
and seasonal effects. The equations that govern this model for
seasonal period, $p$, are given by

\begin{eqnarray*}
a_t &=& \alpha (x_t - s_{t-p}) + (1 - \alpha) (a_{t-1} - b_{t-1}),\\
b_t &=& \beta  (a_t - a_{t-1}) + (1 - \beta)  b_{t-1},\\
s_t &=& \gamma (x_t - a_t)     + (1 - \gamma) s_{t-p},
\end{eqnarray*}

where

\begin{eqnarray*}
a_t && \text{is the estimated level at time $t$},\\
b_t && \text{is the estimated slope at time $t$},\\
s_t && \text{is the estimated seasonal effect at time $t$},\\
\alpha, \beta, \gamma && \text{are smoothing parameters}
\end{eqnarray*}

Within the framework of the `forecast` package, we have an additive trend with
an additive seasonal component and additive errors. This corresponds to using
the 'AAA' `ets()` model


#### Air Passenger Data

We start with the Air Passenger data, and use the 'AAA' model within `ets()`.

```{r fit_holtwinters_ts_model, echo=TRUE}
ap_hw_fit <- ap_train_tbl %>%
  tk_ts(select = value, start = 1949, frequency = 12) %>%
  ets(model = 'AAA')

ap_hw_fit %>% sw_tidy()
ap_hw_fit %>% sw_glance()


ap_hw_forecast <- ap_hw_fit %>%
  forecast(h = 12, level = c(50, 80))

ap_hw_forecast_tbl <- ap_hw_forecast %>%
  sw_sweep(rename_index = 'month')


ggplot(ap_hw_forecast_tbl) +
  geom_line  (aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = ap_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Forecast for Air Passengers with Holt-Winters Model')


ap_hw_augment_tbl <- ap_hw_fit %>%
  sw_augment(rename_index = 'month')

ggplot(ap_hw_augment_tbl) +
  geom_line(aes(x = month, y = .actual)) +
  geom_line(aes(x = month, y = .fitted), colour = 'red') +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('sw_augment Output for Holt-Winters Model')
```

For this data the multiplicative trend and errors is not the most appropriate
so we will refit this data with those changes.

Rather than run through the full set of code again we run a more concise
version of this code.

```{r fit_ap_mult_hw, echo=TRUE}
ap_hwmult_fit <- ap_train_tbl %>%
  tk_ts(select = value, start = 1949, frequency = 12) %>%
  ets(model = 'MAM')

ap_hwmult_forecast <- ap_hwmult_fit %>%
  forecast(h = 12, level = c(50, 80))

ap_hwmult_forecast_tbl <- ap_hwmult_forecast %>%
  sw_sweep(rename_index = 'month')


ap_hwmult_fit %>% sw_tidy()
ap_hwmult_fit %>% sw_glance()

ggplot(ap_hwmult_forecast_tbl) +
  geom_line(aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = ap_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Air Passengers Forecast for Multiplicative Holt-Winters Model')
```


#### Maine Unemployment Data

We now try to fit our models with the Maine Unemployment data, repeating the
process from before.

```{r fit_maine_hw, echo=TRUE}
maine_hw_fit <- maine_train_tbl %>%
  tk_ts(select = value, start = 1996, frequency = 12) %>%
  ets(model = 'ANN')

maine_hw_forecast <- maine_hw_fit %>%
  forecast(h = 20, level = c(50, 80))

maine_hw_forecast_tbl <- maine_hw_forecast %>%
  sw_sweep(rename_index = 'month')


maine_hw_fit %>% sw_tidy()
maine_hw_fit %>% sw_glance()


ggplot(maine_hw_forecast_tbl) +
  geom_line(aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = maine_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Additive Holt-Winters Model for the Maine Unemployment Data')
```

As before, we can try a multiplicative model.

```{r fit_maine_hwmult, echo=TRUE}
maine_hwmult_fit <- maine_train_tbl %>%
  tk_ts(select = value, start = 1996, frequency = 12) %>%
  ets(model = 'MAM')

maine_hwmult_forecast <- maine_hwmult_fit %>%
  forecast(h = 20, level = c(50, 80))

maine_hwmult_forecast_tbl <- maine_hwmult_forecast %>%
  sw_sweep(rename_index = 'month')


maine_hwmult_fit %>% sw_tidy()
maine_hwmult_fit %>% sw_glance()


ggplot(maine_hwmult_forecast_tbl) +
  geom_line(aes(x = month, y = value, colour = key)) +
  geom_ribbon(aes(x = month, ymin = lo.50, ymax = hi.50), alpha = 0.2) +
  geom_line  (aes(x = month, y = value), colour = 'red', data = maine_test_tbl) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Count") +
  scale_x_yearmon() +
  ggtitle('Multiplicative Holt-Winters Model for the Maine Unemployment Data')
```

We see that including a seasonal component into the model improves the
predictions

