---
title: "Time Series Analysis with R"
output: 
  flexdashboard::flex_dashboard:
    theme: united
    css: style.css
---

```{r knit_opts, include = FALSE}
#rm(list = ls()); gc()

knitr::opts_chunk$set(tidy       = FALSE,
                      cache      = FALSE,
                      message    = FALSE,
                      warning    = FALSE,
                      fig.height =     8,
                      fig.width  =    11)

library(conflicted)
library(tibble)
library(tidyverse)
library(dplyr)
library(stringr)
library(magrittr)
library(tidyquant)
library(ggplot2)
library(scales)
library(cowplot)
library(cranlogs)
library(timetk)
library(sweep)
library(seasonal)
library(forecast)
library(broom)

conflict_prefer('select',  'dplyr')
conflict_prefer('filter',  'dplyr')
conflict_prefer('lag',     'dplyr')
conflict_prefer('extract', 'magrittr')



options(width = 80L,
        warn  = 1
        )

set.seed(42)

theme_set(theme_cowplot())


source('custom_functions.R')
```

Introduction {data-navmenu="Introduction"}
============================================
column {.tabset}
---------------------------
### Introduction

<h4>Introduction</h4>

Time series occur in most fields of study that produce quantitative data.
Whenever quantities are measured over time, those measurements form a
time-series, or more formally, a discrete-time stochastic process.


The content of this workshop is partly based on the book "Introductory Time
Series with R" by Paul Cowpertwait and Andrew Metcalfe. 

While available as in dead-tree form, the full text is available online:

https://otexts.com/fpp2/





### R Packages

In this workshop we focus on the use of 'tidy'-style tools in the analysis of
time-series. In particular we look at packages such as `tidyquant` that enable
and simplify this approach to time-series analysis.


Basic Concepts {data-navmenu="Introduction"}
==========================================
column {.tabset}
---------------------------
### Basic Concepts

A famous example of a time-series is count of airline passengers in the US,
as shown in the figure below. This is a simple univariate time-series, with
measurements taken on a monthly basis over a number of years, each datum
consisting of a single number - the number of passengers travelling via
a commercial flight that month.

```{r plot_air_passenger_timeseries, echo=TRUE}
plot(AirPassengers)
```

Before we begin analysing such data, we first need to create a mathematical
framework to work in. Fortunately, we do not need anything too
complicated, and for a finite time-series of length $N$, we model the time
series as a sequence of $N$ random variables, $X_i$, with $i = 1, 2, ..., N$.

Note that each individual $X_i$ is a wholly separate random variable: we only
ever have a single measurement from each random variable. In many cases we
simplify this, but it is important to understand and appreciate that such
simplifications are just that. Time series are difficult to analyse.

Before we get to any of that though, and before we try to build any kind of
models for the data, we always start with visualising the data. Often, a simple
plot of the data helps use pick out aspects to analyse and incorporate into the
models. For time series, one of the first things to do is the *time plot*, a
simple plot of the data over time.

For the passenger data, a few aspects stand out that are very common in time
series. It is apparent that the numbers increase over time, and this systematic
change in the data is called the *trend*. Often, approximating the trend as a
linear function of time is adequate for many data sets.

A repeating pattern in the data that occurs over the period of the data (in
this case, each year), is called the *seasonal variation*, though a more
general concept of 'season' is implied --- it often will not coincide with the
seasons of the calendar.

A slightly more generalised concept from the seasonality is that of *cycles*,
repeating patterns in the data that do not correspond to the natural fixed
periods of the model. None of these are apparent in the air passenger data, and
accounting for them are beyond the scope of this introductory tutorial.

Finally, another important benefit of visualising the data is that it helps
identify possible *outliers* and *erroneous* data.

In many cases, we will also be dealing with time series that have multiple
values at all, many or some of the points in time.

Often, these values will be related in some ways, and we will want to analyse
those relationships also. In fact, one of the most efficient methods of
prediction is to find *leading indicators* for the value or values you wish to
predict --- you can often use the current values of the leading indicators to
make inference on future values of the related quantities.

The fact that this is one of the best methods in time series analysis says a
lot about the difficulty of prediction (Yogi Berra, a US baseball player noted
for his pithy statements, once said "Prediction is difficult, especially about
the future").


Example Timeseries {data-navmenu="Introduction"}
==================================================
column {.tabset}
--------------------------
### Example Data Sets 
In this workshop we will look at a number of different time-series, discussed
here.

This data comes in a few different format, and this workshop discusses methods
for analysing this data in a common format.


### Air Passenger Data

As mentioned previously, a canonical time-series is the airline passenger
dataset, and this is the first dataset we look at.

```{r show_airline_passengers}
# AirPassengers %>% summary()

AirPassengers %>% plot()
```

In this workshop we will convert all time series into the tibbles: the package
`timetk` allows us to do this.

```{r convert_airpassengers_tibble, echo=TRUE}
airpassengers_tbl <- AirPassengers %>% tk_tbl(rename_index = 'month')
```

```{r}
#airpassengers_tbl %>% head()
```

```{r}

ggplot(airpassengers_tbl) +
  geom_line(aes(x = month, y = value)) +
  expand_limits(y = 0) +
  xlab('Date') +
  ylab('Passenger Count') +
  scale_x_yearmon() +
  ggtitle('Plot of Air Passenger Time Series')
```


### Maine Unemployment Data

Time series are very common in econometrics, and a dataset provided in the
text is that of monthly unemployment statistics in Maine from 1996 on. I have
included the datafile in this workshop.

```{r load_maine_unemployment, echo=TRUE}
maine_ts <- scan('data/Maine.dat', skip = 1) %>%
  ts(start = 1996, frequency = 12)

#maine_ts %>% summary()
maine_ts %>% plot()
```

As before, we convert this data into a tibble and recreate the plot using
`ggplot2`.

```{r maine_unemployment_tibble, echo=TRUE}
maine_tbl <- maine_ts %>% tk_tbl(rename_index = 'month')

#maine_tbl %>% summary()
```

```{r}

ggplot(maine_tbl) +
  geom_line(aes(x = month, y = value)) +
  expand_limits(y = 0) +
  xlab('Date') +
  ylab('Unemployment Numbers') +
  scale_x_yearmon() +
  ggtitle('Plot of Maine Unemployment Time Series')
```


### Australian Consumption Statistics (CBE)

Governments produce regular data on consumption numbers for their economy.

One
such dataset is contained in the file `cbe.dat`, produced by the Australian
Census Bureau containing data of chocolate, beer and energy production on a
monthly basis.

```{r load_australian_cbe_data, echo=TRUE}
library(readr)
cbe_raw_tbl <- read_tsv('data/cbe.dat')

#cbe_raw_tbl %>% glimpse()
```

Similar to the Maine file, this data does not contain time indices for the
data. For the sake of completeness, we use the same approach as before and
convert to a tibble, but we will also show how to construct the time index
without having to do intermediate conversions.

First we add time indices via `ts` conversions.


```{r construct_cbe_timeseries, echo=TRUE}
cbe_ts <- cbe_raw_tbl %>%
  as.matrix() %>%
  ts(start = 1958, frequency = 12)

cbe_ts %>% head(10)
cbe_ts %>% plot()
```

An alternative approach is to add the time index directly.

```{r australian_cbe_add_time, echo=TRUE}
n_data <- cbe_raw_tbl %>% nrow()

cbe_tbl <- cbe_raw_tbl %>%
  add_column(month = (1958 + (0:(n_data-1)/12)) %>% yearmon, .before = 1)


cbe_tbl %>% glimpse()
#cbe_tbl %>% summary()
```

Having constructed the tibble, we now construct these time series plots using
`ggplot2`.

```{r construct_cbe_plots, echo=TRUE}
plot_tbl <- cbe_tbl %>%
  gather('product', 'value', -month)

ggplot(plot_tbl) +
  geom_line(aes(x = month, y = value, colour = product)) +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Production Amount") +
  scale_x_yearmon() +
  scale_y_continuous(labels = comma) +
  ggtitle('Production Data from Australian Government')
```


Due to the different scales, it might be more useful to use a faceted plot for
each product:

```{r plot_cbe_facetted_data, echo=TRUE}
plot_tbl <- cbe_tbl %>%
  gather('product', 'value', -month)

ggplot(plot_tbl) +
  geom_line(aes(x = month, y = value)) +
  facet_grid(rows = vars(product), scales = 'free_y') +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Production Amount") +
  scale_x_yearmon() +
  scale_y_continuous(labels = comma) +
  ggtitle('Production Data from Australian Government')
```


### Australian Building Activity

Another set of data produced by the Australian Census Bureau is the Building
Activity Publication listing the value of building work done in each quarter.
This data contains the total dwellings approved for construction per month,
averaged over the past three months, and the value of work done over the past
three months.

This data is quarterly from 1996, and we construct the data as before.


```{r read_approvactiv_data, echo=TRUE}
approvactiv_raw_tbl <- read_tsv('data/ApprovActiv.dat')

approvactiv_raw_tbl %>% glimpse()


approvactiv_ts <- approvactiv_raw_tbl %>%
  as.matrix() %>%
  ts(start = 1996, frequency = 4)

approvactiv_ts %>% head(10)
approvactiv_ts %>% plot()
```
```{r}

approvactiv_tbl <- approvactiv_ts %>%
  tk_tbl(rename_index = 'quarter')

approvactiv_tbl %>% glimpse()
#approvactiv_tbl %>% summary()
```

As before, we can produce time plots in `ggplot2`.


```{r construct_approvactiv_plots, echo=TRUE}
plot_tbl <- approvactiv_tbl %>%
  gather('label', 'value', -quarter)

ggplot(plot_tbl) +
  geom_line(aes(x = quarter, y = value)) +
  facet_grid(rows = vars(label), scales = 'free_y') +
  expand_limits(y = 0) +
  xlab("Date") +
  ylab("Amount") +
  scale_x_yearqtr() +
  scale_y_continuous(labels = comma) +
  ggtitle('Building Activity Data from Australian Government')
```



### CRAN Package Downloads Data

An interesting dataset is the daily count of package downloads from CRAN. This
data is easy to obtain via use of the package `cranlogs`, which gives us use
of the `cran_downloads()` function.

For this workshop, we will look at some of the main packages that comprise
the 'tidyverse', as well as the total number of downloads from CRAN.

```{r download_cran_data, echo=TRUE}
cran_data_file <- 'data/cran_download_data.csv'

if(file.exists(cran_data_file)) {
  cran_data_tbl <- read_csv(cran_data_file)
} else {
  cran_pkgs <- c('dplyr', 'tidyr', 'ggplot2', 'lubridate', 'stringr', 'tibble',
                 'broom', 'jsonlite', 'purrr', 'readr', 'tidyquant')
    

  cran_data_tbl <- retrieve_cran_download_data(cran_pkgs, '2014-01-01', '2018-12-31')
  
  cran_data_tbl %>% write_csv(path = cran_data_file)
}


cran_data_tbl %>% glimpse()
#cran_data_tbl %>% summary()
```

First we construct a simple line plot of the download counts, facetted by
package.

```{r plot_cran_downloads, echo=TRUE}
ggplot(cran_data_tbl) +
  geom_line(aes(x = date, y = count)) +
  facet_wrap(vars(package), scales = 'free_y') +
  expand_limits(y = 0) +
  scale_y_continuous(labels = comma) +
  ggtitle('Facetted Lineplots of CRAN Daily Downloads')
```

Not all packages have download data as some packages were created after the
start of our observation period. This manifests as zero counts for that
package. We discuss these issues later on in the workshop.


Combining Time Series {data-navmenu="Introduction"}
==========================================
column {.tabset}
---------------------------------------
Useful insights are often gained from combining different datasets together.

Looking at our datasets, one possible interesting relationship is that between
energy production and airline passengers -- it is reasonable to expect that
both of these quantities will be related as they are related to the overall
health and size of the economy.

A major benefit of using tidy tools for time series is to make such data
manipulation and arrangement simple: combining datasets is simply a matter
of using the two-table joins.

To illustrate, we combine the Air Passenger data with the Australian economic
data, using the following code. Note that we rename the air passenger data
at the end to make it more meaningful and useful.


```{r combining_airpassengers_australia_data, echo=TRUE}
ap_econ_combined_tbl <- airpassengers_tbl %>%
  left_join(cbe_tbl, by = 'month') %>%
  filter(complete.cases(.)) %>%
  rename(air = value)


ap_econ_combined_tbl %>% glimpse()
#ap_econ_combined_tbl %>% summary()
```

We return to this dataset later in the workshop.


# Manipulation of Time Series Data

Much like all data, it is rate to get time-series exactly in the format we
want for analysis. For various reasons, we may want to analyse transformations
or aggregations of this data.

Much like feature engineering and variable selection, this process can be
more art than science - there are no hard and fast rules, merely principles
and rules-of-thumb.

The last few years in particular have seen the development of a number of
tools to aid us with the analysis of time series along 'tidy' principles. In
particular, we will focus on the use of `tidyquant` - a package aimed at
analysing financial data, but which is also very useful for time series.


## Aggregating Data {data-navmenu="Introduction"}
==========================================
column {.tabset}
---------------------------------------

From a conceptual point of view, aggregating time series is the most
straightforward - we group the data by longer periods of time and aggregate
each individual 'bucket' of data as desired or required.


### Single Statistics

As an example of this, suppose we wish to look at the air passenger data as an
annual sum for each year. Our data is monthly, so we need to aggregate this
data into annual numbers.

```{r aggregate_ap_data_annual, echo=TRUE}
ap_yearly_dplyr_tbl <- airpassengers_tbl %>%
  group_by(year = month %>% format('%Y') %>% as.numeric()) %>%
  summarise(ann_total = sum(value))


ap_yearly_dplyr_tbl %>% glimpse()
#ap_yearly_dplyr_tbl %>% summary()
```

The above transformation was straightforward using existing `dplyr`
functionality, but we can also use routines provided for by `tidyquant` and its
function `tq_transmute`:


```{r aggregate_ap_annual_transmute, echo=TRUE}
ap_yearly_tidyquant_tbl <- airpassengers_tbl %>%
  tq_transmute(
    select     = value,
    mutate_fun = apply.yearly,
    FUN        = sum,
    na.rm      = TRUE,
    col_rename = 'ann_total'
  ) %>%
  mutate(year = month %>% format('%Y') %>% as.numeric())


ap_yearly_tidyquant_tbl %>% glimpse()
#ap_yearly_tidyquant_tbl %>% summary()
```

We can now look at a lineplot for this new time-series of annual totals. This
gives us an idea of the overall trend in the data.


```{r aggregate_ap_data_annual_lineplot, echo=TRUE}
ggplot(ap_yearly_tidyquant_tbl) +
  geom_line(aes(x = year, y = ann_total)) +
  expand_limits(y = 0) +
  scale_y_continuous(labels = comma) +
  ggtitle('Lineplot of the Annual Air Passenger Totals')
```

#### Exercises

  1. Aggregate the Maine unemployment data into yearly totals.
  1. Create the relevant lineplots to check this data for trends and patterns.
  1. Aggregate the CBE data into yearly totals.
  1. Create plots of this data to check for trends.


### Multiple Statistics

Should we need multiple statistics in our output, we implement this by
writing a custom function that outputs the statistics we require.

```{r aggregate_ap_annual_transmute_custom, echo=TRUE}
custom_stats_func <- function(x, na.rm = TRUE, ...) {
  c(
    sum    = sum(x, na.rm = na.rm),
    mean   = mean(x, na.rm = na.rm),
    sd     = sd(x, na.rm = na.rm),
    median = median(x, na.rm = na.rm),
    q25    = quantile(x, na.rm = na.rm, probs = 0.25, names = FALSE),
    q75    = quantile(x, na.rm = na.rm, probs = 0.75, names = FALSE)
  )
}


ap_yearly_tidyquant_custom_tbl <- airpassengers_tbl %>%
  tq_transmute(
    select     = value,
    mutate_fun = apply.yearly,
    FUN        = custom_stats_func,
    na.rm      = TRUE
  ) %>%
  mutate(year = month %>% format('%Y') %>% as.numeric())


ap_yearly_tidyquant_custom_tbl %>% glimpse()
# ap_yearly_tidyquant_custom_tbl %>% summary()
```


#### Exercises

  1. Repeat the aggregation shown using the Maine unemployment data
  1. Repeat the aggregation shown using the CBE data.
  1. Create a custom function to calculate mean, sd, skew and kurtosis


## Rolling Functions {data-navmenu="Introduction"}
==========================================
column {.tabset}
---------------------------------------

Another common transformation of time-series is to apply a function over a
fixed rolling window of data.

Note that rolling functions different conceptually from aggregates as they
are not calculated over disjoint subsets of the data: the output is at the
same time period as the original data.

Because of this difference we use a different function from `tidyquant` to
execute this calculation: `tq_mutate()`:


### Moving Averages

A common rolling function is the *moving average*: we calculate the average
value of the time series over a fixed window of data.

```{r rolling_ap_sixmonth_mutate, echo=TRUE}
ap_rollmean_sixmonth_tbl <- airpassengers_tbl %>%
  tq_mutate(
    # tq_mutate args
    select     = value,
    mutate_fun = rollapply, 
    # rollapply args
    width      = 6,
    align      = "right",
    FUN        = mean,
    # mean args
    na.rm      = TRUE,
    # tq_mutate args
    col_rename = "mean_6m"
  )


ap_rollmean_sixmonth_tbl %>% glimpse()
#ap_rollmean_sixmonth_tbl %>% summary()
```

We compare the two values by plotting the original time series against its
moving average.


```{r rolling_ap_sixmonth_plot, echo=TRUE}
plot_tbl <- ap_rollmean_sixmonth_tbl %>%
  rename(orig = value) %>%
  gather('label', 'value', -month)


ggplot(plot_tbl) +
  geom_line(aes(x = month, y = value, colour = label)) +
  expand_limits(y = 0) +
  xlab('Month') +
  ylab('Passenger Total') +
  scale_x_yearmon() +
  scale_y_continuous(labels = comma) +
  ggtitle('Comparison Plot of the Air Passenger Counts')
```

Note that the moving-average series does not start at the same timestamp as
the original dataset size is reduced by the windowing function.

We can add multiple moving averages to a time series by chaining a series of
`tq_mutate()` calls together.


```{r rolling_ap_multiwindow, echo=TRUE}
ap_rollmean_multi_tbl <- airpassengers_tbl %>%
  tq_mutate(
    # tq_mutate args
    select     = value,
    mutate_fun = rollapply,
    # rollapply args
    width      = 6,
    align      = "right",
    FUN        = mean,
    # mean args
    na.rm      = TRUE,
    # tq_mutate args
    col_rename = "mean_6m"
  ) %>%
  tq_mutate(
    # tq_mutate args
    select     = value,
    mutate_fun = rollapply, 
    # rollapply args
    width      = 12,
    align      = "right",
    FUN        = mean,
    # mean args
    na.rm      = TRUE,
    # tq_mutate args
    col_rename = "mean_12m"
  )
  

ap_rollmean_multi_tbl %>% glimpse()
ap_rollmean_multi_tbl %>% summary()
```

As before, we now create a lineplot of the three values to show the effect of
the different window sizes.

```{r rolling_ap_multi_plot, echo=TRUE}
plot_tbl <- ap_rollmean_multi_tbl %>%
  rename(orig = value) %>%
  gather('label', 'value', -month)


ggplot(plot_tbl) +
  geom_line(aes(x = month, y = value, colour = label)) +
  expand_limits(y = 0) +
  xlab('Month') +
  ylab('Passenger Total') +
  scale_x_yearmon() +
  scale_y_continuous(labels = comma) +
  ggtitle('Comparison Plot of the Air Passenger Counts')
```

The twelve month time series is shorter than the six month series as it has
a wider calculation window.

Any sort of other windowing functions can be applied, including the standard
deviation, allowing us to include a range of possible values.


```{r rolling_ap_ribbon, echo=TRUE}
ribbon_func <- function(x, na.rm = TRUE, ...) {
  mu    <- mean(x, na.rm = na.rm)
  sigma <- sd(x, na.rm = na.rm)
  
  lower <- mu - 2 * sigma
  upper <- mu + 2 * sigma
  
  return(c(mu = mu, l2sd = lower, u2sd = upper))
}


ap_roll_ribbon_tbl <- airpassengers_tbl %>%
  tq_mutate(
    # tq_mutate args
    select     = value,
    mutate_fun = rollapply, 
    # rollapply args
    width      = 6,
    align      = "right",
    by.column  = FALSE,
    FUN        = ribbon_func,
    # mean args
    na.rm      = TRUE
  )
  

ap_roll_ribbon_tbl %>% glimpse()
ap_roll_ribbon_tbl %>% summary()
```

We now plot the original data against the moving average and the mean.


```{r plot_ap_6m_ribbon_data, echo=TRUE}
ggplot(ap_roll_ribbon_tbl) +
  geom_line(aes(x = month, y = value)) +
  geom_line(aes(x = month, y = mu), colour = 'red') +
  geom_ribbon(aes(x = month, ymin = l2sd, ymax = u2sd),
              colour = 'grey', alpha = 0.25) +
  expand_limits(y = 0) +
  xlab('Month') +
  ylab('Passenger Total') +
  scale_x_yearmon() +
  scale_y_continuous(labels = comma) +
  ggtitle('Ribbon Plot of the Air Passenger Counts (6 month window)')
```


We now repeat this process with using a twelve-month window for the data.

```{r rolling_ap_12m_ribbon, echo=TRUE}
ap_roll_12m_ribbon_tbl <- airpassengers_tbl %>%
  tq_mutate(
    # tq_mutate args
    select     = value,
    mutate_fun = rollapply, 
    # rollapply args
    width      = 12,
    align      = "right",
    by.column  = FALSE,
    FUN        = ribbon_func,
    # mean args
    na.rm      = TRUE
  )
  

ap_roll_12m_ribbon_tbl %>% glimpse()
ap_roll_12m_ribbon_tbl %>% summary()
```

Having constructed the data, we once again create a ribbon plot with these
quantities.

```{r plot_ap_12m_ribbon_data, echo=TRUE}
ggplot(ap_roll_12m_ribbon_tbl) +
  geom_line(aes(x = month, y = value)) +
  geom_line(aes(x = month, y = mu), colour = 'red') +
  geom_ribbon(aes(x = month, ymin = l2sd, ymax = u2sd),
              colour = 'grey', alpha = 0.25) +
  expand_limits(y = 0) +
  xlab('Month') +
  ylab('Passenger Total') +
  scale_x_yearmon() +
  scale_y_continuous(labels = comma) +
  ggtitle('Ribbon Plot of the Air Passenger Counts (12 month window)')
```


#### Exercises

  1. Construct a 3 month moving average for the passenger data and compare it
     to the 6 and 12 month values.
  1. Calculate the 6 month and 12 month rolling average values for the Maine
     unemployment data.
  1. Construct the ribbon plot for the Maine unemployment data.
  1. Construct moving average data for the CBE dataset. This process may be
     made easier by reshaping the data.


### Differences

Another common transformation of the data is to take the 'first differences'
of the values, i.e. we convert the time series of values into one of
differences. We discuss the reasons for this later on -- for now we focus on
the mechanics of creating first differences.


```{r ap_firstdiff_mutate, echo=TRUE}
ap_firstdiff_tbl <- airpassengers_tbl %>%
  mutate(diff = value - lag(value, n = 1))


ap_firstdiff_tbl %>% glimpse()
ap_firstdiff_tbl %>% summary()
```

Having calculated the differences, we now produce a lineplot of those values.

```{r ap_firstdiff_plot, echo=TRUE}
plot_tbl <- ap_firstdiff_tbl %>%
  rename(count = value) %>%
  gather('series', 'value', -month)


ggplot(plot_tbl) +
  geom_line(aes(x = month, y = value, colour = series)) +
  expand_limits(y = 0) +
  xlab('Month') +
  ylab('Value') +
  scale_x_yearmon() +
  scale_y_continuous(labels = comma) +
  ggtitle('Plot of the Air Passenger Counts and First Differences')
```

As we see with this plot, the first differences of the passenger data does not
contain a trend.


#### Exercises

  1. Calculate the first differences for the Maine unemployment data.
  1. Create a lineplot of this data to check for its value.
  1. Calculate the first differences for the CBE data.
  1. Create lineplots for the CBE differences.
  1. Using the `lag()` function with the Air Passenger data, calculate the
     percentage changes data instead of the arithmetic changes.
  1. Construct the lineplot for the percentage change values.
  
  

# Exploratory Data Analysis of Time Series

The first step in all exploratory analysis is simple visualisation: simple
lines plots such as those we have seen are our starting point. The human
brain is excellent at pattern recognition, so a simple plot often guides our
analysis more effectively than a suite of numerical diagnostics.

Rather than discuss different techniques, we will explore our sample data as a
way to illustrate some ways of initially exploring the datasets.


## Air Passenger {data-navmenu="Case Studies"}
==========================================
column {.tabset}
---------------------------------------

We start with the air passenger data, and remind ourselves of the basic
structure of the data.

```{r airpassengers_explore_struct, echo=TRUE}
airpassengers_tbl %>% glimpse()
airpassengers_tbl %>% summary()
```


### Raw Data

We remind ourselves of the lineplot of the raw data.

```{r plot_airpassenger_exploration, echo=TRUE}
ggplot(airpassengers_tbl) +
  geom_line(aes(x = month, y = value)) +
  xlab('Date') +
  ylab('Passenger Count') +
  expand_limits(y = 0) +
  scale_x_yearmon() +
  ggtitle('Plot of Air Passenger Time Series')
```

This plot suggests a strong seasonal effect as well as a trend so this is the
first thing to investigate.

It may help if we can add points to the plot to indicate the months, so we add
points for the month of June to help us identify years.

```{r plot_airpassenger_exploration_dec_points, echo=TRUE}
ap_jun_tbl <- airpassengers_tbl %>% filter(format(month, '%m') == '06')

ggplot(airpassengers_tbl) +
  geom_line(aes(x = month, y = value)) +
  geom_point(aes(x = month, y = value), data = ap_jun_tbl, size = 2) +
  expand_limits(y = 0) +
  xlab('Date') +
  ylab('Passenger Count') +
  scale_x_yearmon() +
  ggtitle('Plot of Air Passenger Time Series')
```



To look into trends, we have a number of options: we could look at yearly
sums or averages, or we could look at moving averages.

```{r plot_ap_yearly_totals, echo=TRUE}
ggplot(ap_yearly_tidyquant_tbl) +
  geom_line(aes(x = year, y = ann_total)) +
  expand_limits(y = 0) +
  scale_x_yearmon() +
  scale_y_continuous(labels = comma) +
  ggtitle('Lineplot of the Annual Air Passenger Totals')
```




One way to investigate the seasonality in the dataset is to construct a boxplot
of the passenger counts, grouping by month.

```{r plot_ap_monthly_boxplots, echo=TRUE}
plot_tbl <- airpassengers_tbl %>%
  mutate(cal_month = format(month, '%m'))

ggplot(plot_tbl) +
  geom_boxplot(aes(x = cal_month, y = value)) +
  expand_limits(y = 0) +
  xlab('Month') +
  ylab('Passenger Count') +
  scale_x_yearmon() +
  scale_y_continuous(labels = comma) +
  ggtitle('Boxplot of the Air Passenger Counts')
```

We can see some aspects of the data seasonality in this boxplot, but the
multiplicative nature of the plots means the seasonal trends is obscured a
little.

### First Differences

We also produce a similar boxplot, but this time looking at the first
differences.

```{r plot_ap_diffs_monthly_boxplots, echo=TRUE}
plot_tbl <- ap_firstdiff_tbl %>%
  mutate(cal_month = format(month, '%m'))

ggplot(plot_tbl) +
  geom_boxplot(aes(x = cal_month, y = diff)) +
  expand_limits(y = 0) +
  xlab('Month') +
  ylab('Passenger Count Changes') +
  scale_x_yearmon() +
  scale_y_continuous(labels = comma) +
  ggtitle('Boxplot of the First Differences of the Air Passenger Counts')
```

The seasonality in the data comes through much stronger with this plot. We see
much bigger monthly effects.


### Percentage Changes

To look into multiplicative effects we check the percentage changes from month
to month.


```{r ap_perc_change_explore_struct, echo=TRUE}
ap_perc_change_tbl <- airpassengers_tbl %>%
  mutate(cal_month   = format(month, '%m'),
         perc_change = value / lag(value, n = 1) - 1)


ap_perc_change_tbl %>% glimpse()
ap_perc_change_tbl %>% summary()
```

Having looked at the data and the column, we now look at some simple lineplots
as before.

```{r plot_ap_perc_change_lineplot, echo=TRUE}
ggplot(ap_perc_change_tbl) +
  geom_line(aes(x = month, y = perc_change)) +
  expand_limits(y = 0) +
  xlab('Date') +
  ylab('Percentage Change') +
  scale_x_yearmon() +
  ggtitle('Lineplot of the Percentage Changes of the Air Passenger Counts')
```

Much like the arithmetic differences, the percentage changes are centred around
zero, so now we can look at a boxplot of them.

```{r plot_ap_perc_changes_monthly_boxplots, echo=TRUE}
ggplot(ap_perc_change_tbl) +
  geom_boxplot(aes(x = cal_month, y = perc_change)) +
  xlab('Month') +
  ylab('Percentage Changes') +
  scale_x_yearmon() +
  ggtitle('Boxplot of the Percentage Changes of the Air Passenger Counts')
```


## Maine Unemployment {data-navmenu="Case Studies"}
==========================================
column {.tabset}
---------------------------------------

We now explore the Maine unemployment data.

As before, we look at the raw data, the first differences and the percentage
changes.

```{r maine_data_exploration, echo=TRUE}
maine_explore_tbl <- maine_tbl %>%
  mutate(cal_month   = format(month, '%m'),
         diff        = value - lag(value, n = 1),
         perc_change = value / lag(value, n = 1) - 1
         )


maine_explore_tbl %>% glimpse()
maine_explore_tbl %>% summary()
```


### Raw Data

We start by the standard lineplot of the values.

```{r plot_maine_lineplot, echo=TRUE}
ggplot(maine_explore_tbl) +
  geom_line(aes(x = month, y = value)) +
  expand_limits(y = 0) +
  xlab('Time') +
  ylab('Maine Unemployment Count (thousands)') +
  scale_x_yearmon() +
  scale_y_continuous(labels = comma) +
  ggtitle('Lineplot of the Maine Unemployment Data')
```

We do not see a trend in this data but the data does seem to have strong
seasonal patterns.

To investigate the seasonality we construct the monthly boxplots from the raw
data. Employment is often seasonal in nature - for example, retail is very
busy towards Christmas each year. As such, we expect to see a large seasonal
component in this data.

```{r plot_maine_monthly_boxplot, echo=TRUE}
ggplot(maine_explore_tbl) +
  geom_boxplot(aes(x = cal_month, y = value)) +
  expand_limits(y = 0) +
  xlab('Month') +
  ylab('Maine Unemployment Count (thousands)') +
  scale_x_yearmon() +
  scale_y_continuous(labels = comma) +
  ggtitle('Monthly Boxplot of the Maine Unemployment Data')
```

We see definite changes over the months, though the effect does not seem as
pronounced here as it was for the airline passenger counts.

### First Differences

We now look at first differences for the unemployment data, and start with the
line plot.

```{r plot_maine_diff_lineplot, echo=TRUE}
ggplot(maine_explore_tbl) +
  geom_line(aes(x = month, y = diff)) +
  expand_limits(y = 0) +
  xlab('Time') +
  ylab('Maine Unemployment Count Differences') +
  scale_x_yearmon() +
  ggtitle('Lineplot of the Differences in Maine Unemployment Data')
```

The original data does not exhibit a strong trend, and the first differences
are similar.

```{r plot_maine_diff_monthly_boxplot, echo=TRUE}
ggplot(maine_explore_tbl) +
  geom_boxplot(aes(x = cal_month, y = diff)) +
  expand_limits(y = 0) +
  xlab('Month') +
  ylab('Maine Unemployment Count Differences') +
  scale_x_yearmon() +
  ggtitle('Monthly Boxplot of the Maine Unemployment Data')
```

We see strong seasonal differences in the monthly data.


### Percentage Changes

We look at the lineplot of the percentage changes next.

```{r plot_maine_perc_change_lineplot, echo=TRUE}
ggplot(maine_explore_tbl) +
  geom_line(aes(x = month, y = perc_change)) +
  expand_limits(y = 0) +
  xlab('Time') +
  ylab('Maine Unemployment Count Percentage Changes') +
  scale_x_yearmon() +
  ggtitle('Lineplot of the Percentage Changes in Maine Unemployment Data')
```

As for the differences, no major trends or patterns emerge from this plot.

We now look for seasonal patterns in the percentage changes.

```{r plot_maine_perc_change_monthly_boxplot, echo=TRUE}
ggplot(maine_explore_tbl) +
  geom_boxplot(aes(x = cal_month, y = perc_change)) +
  expand_limits(y = 0) +
  xlab('Month') +
  ylab('Maine Unemployment Count Percentage Changes') +
  scale_x_yearmon() +
  ggtitle('Monthly Boxplot of the Percentage Changes in Maine Unemployment Data')
```



# Time Series Decomposition

Many time series are dominated by trends or seasonal effects, and we can
create fairly simple models based on these two components. The first of these,
the *additive decompositional model*, is just the sum of these effects, with
the residual component being treated as random:

$$
x_t = m_t + s_t + z_t,
$$

where, at any given time $t$,

\begin{eqnarray*}
x_t && \text{is the observed value} \\
m_t && \text{is the trend} \\
s_t && \text{is the seasonal component} \\
z_t && \text{is the error term}
\end{eqnarray*}


It is worth noting that, in general, the error terms will be a correlated
sequence of values, something we will account for and model later.

In other cases, we could have a situation where the seasonal effect
increases as the trend increases, modeling the values as:

$$
x_t = m_t s_t + z_t.
$$

Other options also exist, such as modeling the log of the observed values,
which does cause some non-trivial modeling issues, such as biasing any
predicted values for the time series.

Various methods are used for estimating the trend, such as taking a
*moving average* of the values, which is a common approach.


