###
#### Worksheet Exercise 6.4
###


The provided R code performs the following:

### Code Explanation

1. **Set Random Seed (`set.seed(42);`)**  
   - Ensures reproducibility by initializing the random number generator with a specific seed (42 in this case).

2. **Simulate an AR(1) Time Series (`arima.sim(...)`)**  
   - Simulates a time series using an Autoregressive (AR) model of order 1:
     - `order = c(1, 0, 0)` indicates an AR(1) model with no differencing or moving average terms.
     - `ar = 0.5` specifies the AR coefficient (\( \phi = 0.5 \)), meaning each value in the series depends partly on the previous value with a coefficient of 0.5.
     - `n = 1000` generates 1,000 observations for the time series.
     - `start.innov = 1` initializes the starting innovation (the random shock) as 1.
     - `n.start = 1` uses one starting observation.

   The simulated series (`ar1.ts`) will contain 1,000 values based on these parameters.

3. **Plot the Simulated Series (`plot(ar1.ts);`)**  
   - Visualizes the generated time series. The plot shows:
     - The x-axis: Time indices (1 to 1,000).
     - The y-axis: Values of the simulated AR(1) process.
   - Given \( \phi = 0.5 \), the series will exhibit autocorrelation—each value is partially dependent on its predecessor. Additionally, the random shocks introduce variation around the trend.

The result is a time series plot displaying how the AR(1) model evolves over the specified number of observations. The pattern will appear smooth with dependency between consecutive points but with a stochastic component. Let me know if you'd like further insights into interpreting the plot!
```{r}

set.seed(42);

ar1.ts <- arima.sim(list(order = c(1, 0, 0), ar = 0.5), n = 1000, start.innov = 1, n.start = 1);
```

```{r}
plot(ar1.ts);
```
